Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory

### 2020-10-31 papers/thesis-kirstie/thesis.tex
edits

The goal of this project is to generate computationally temperature-density 
and pressure-temperature phase diagrams for a Weeks-Chandler-Anderson (WCA) 
model fluid using classical Density Functional Theory (cDFT). 
To do this we need to find the Helmholtz free energies that the fluid
would have in a liquid state, and in a crystalline solid state 
at a given temperature and density. 
%Once the Helmholtz free energy for both the crystal and liquid states of 
%the fluid are found for a given temperature and density, they can be 
%compared to find the resulting state of the fluid. 
%
%save:
%Outside the region of 
%coexistence which exists between the liquid and crystal states, and which 
%consists of various amounts of liquid and crystal existing together,
%the fluid will be completely in a crystalline state if the 
%crystal free energy is lower than the liquid free energy, and it will be 
%completely in a liquid state if the liquid free energy is lower than 
%the crystal free energy. Maxwell's Double Tangent Construction is used 
%to determine the actual number densities of the crystal and liquid states
%at the boundaries of the coexistance region for a given temperature from 
%which the phase diagrams can be constructed.
%end save
%
%In general, if at some number density and temperature the crystal free energy 
%is lower than the homogeneous free energy of the liquid, then the fluid will 
%be in the crystalline state. If the homogeneous liquid free energy is lower 
%than the crystal free energy, then the fluid will be in the liquid state. 
%
%To find the equilibrium Helmholtz free energies we use classical Density 
%Functional Theory (cDFT) in conjunction with Soft Fundamental Measure Theory (SFMT).
%%the corresponding equilibrium number density profile. 
%In particular, 
In general, the Helmholtz free energy, 
which consists of an ideal part and an excess part ,
can be written as a functional of 
the number density profile $n(\vec{r})$ which describes the way the average number 
density varies spatially. 
\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
%

Classical Density Functional Theory is used to find the equilibrium 
Helmholtz free energy, and Soft Fundamental Measure Theory (SFMT) will 
be used to form the functional for the excess Helmholtz free energy. 
%to predict the freezing of a Weeks-Chandler-Anderson 
%model fluid by applying SFMT with Schmidt's proposed erf model, and our temperature
%dependent parameters to form the functional for the excess Helmholtz free energy. 
%for generating the weight funcions, and Tarazona's tensor functional. 
%
%Classical Density Functional Theory (cDFT) is used to find the equilbrium 
According to classical Density Functional Theory the equilbrium 
Helmholtz free energy of the system 
can be found
by varying the number density profile $n(\vec{r})$ until the free energy 
is minimized. 
%The minimum free energy, and its corresponding number 
%density profile are then the equilibrium values of the Helmholtz free energy
%of the system at the given temperature and number density. 
%
%To find the Helmholtz free energy both the ideal Helmholtz free energy
%and the excess Helmholtz free energy must be found and added together.
%The ideal Helmholtz free energy is the free energy the fluid would have in
%the absence of any interactomic potential energy $V(r)$ between the spheres, 
%or atoms that make up the fluid. 
%
%The excess Helmholtz free energy is that 
%additional energy beyond the ideal Helmholtz free energy which results from the presence of an interatomic potential 
%energy $V(r)$ between the spheres.
%
%\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
%For a spatially varying number density, as would be expected for a crystal, 
%the free energy is a functional of the number density profile $n(\vec{r})$  
%The Helmholtz free energy excess free energy is then added to the ideal free energy calculated 
%analytically to find the resulting free energy for the system. 
%
SFMT makes the task of forming an expression
%is a method used to form an expression 
for the excess Helmholtz free energy of a particular model fluid 
much easier since the interaction that each sphere, or atom, 
has with every other sphere does not have to be taken explicitly into account.
%without 
%having to take explicitly into account the interaction each sphere, or atom, 
%has with every other sphere which would make the task very difficult, 
%or even prohibtive.
%
%Classically, fluids have been modeled by a collection of small, 
%impenetrable spheres colliding with one another like billiard balls. 
%In the simplest case, the spheres are treated as having only a replulsive 
%potential. This model, called the Hard-Sphere fluid, has served as the 
%primary reference system for fluids. This is due to the fact that the 
%repulsion between atoms is much stronger than their attraction, and so 
%the repulsive potential, which is descriptive of hard spheres, dominates 
%the interactions between atoms. In 1954 Zwanzig introduced the Square-well 
%fluid which is a perturbation of the Hard-Sphere fluid created by adding 
%a hard-edge attractive potential to the repulsive potential of the 
%Hard-Sphere fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied 
%spheres with softly fading replusive potentials which he modeled by making
%the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
%This reflects the idea that soft-spheres penetrate more deeply into 
%each other when they collide at higher temperatures. 
%
%In 1967 Barker and Henderson combined the perturabtion approach of Zwanzig 
%with Rowlinson's idea of using temperature-dependent radii to model spheres 
%with soft repulsive potentials to create the Barker-Henderson model 
%fluid~\cite{barker1967perturbation}, where a soft-sphere fluid by is 
%modeled by using a Hard-Sphere fluid with hard spheres of larger radii 
%at low temperatures, and smaller radii at high temperatures. In 1971 Weeks, 
%Chandler, and Anderson used temperature and density dependent hard-sphere 
%radii to model spheres with the more realistic softly fading repulsive 
%portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
%which is known as the Weeks-Chandler-Anderson model fluid. f
%
%While models for particular fluids were being created, powerful methods 
%to describe the properties of fluids in general were also being developed. 
%In the late 1950s, Scaled Particle Theory (SPT)~\cite{ReissSPT} was 
%introduced from which an equation of state, and an expression for the 
%excess Helmholtz free energy of a homogeneous Hard-Sphere fluid can be 
%obtained~\cite{Hansen}. \color{red}ADD Percus-Yevick introduced an 
%alternative method...OZ... \color{black}
%%Percus and Yevick took another approach to the Hard-Sphere fluid.
%In 1961 Morita and Hiroike introduced a density functional 
%approach~\cite{MoritaDFT} to finding the equilibrium grand free energy 
%(the equilibrium Helmholtz free energy in the case of a canonical 
%ensemble), and the cooresponding equilibrium number density profile for 
%inhomogeneous fluids. To use this approach, a functional must be 
%constructed for a particular fluid.
%%DFT Evans 1979
%%Led by the consistency bewteen results from SPT and Percus-Yevick, 
%In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
%(which can be constructed from SPT~\cite{santos2012phi3}) to form a 
%functional for the Hard-Sphere fluid. Tarazona later improved on the 
%method by introducing a tensor version of the functional~\cite{tarazonaphi3, 
%santos2012phi3}, and in 1999 Schmidt adopted FMT for soft spheres 
%forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}. 
%
%Schmidt applied his Soft Fundamental Measure Theory to various fluids 
%with soft repulsive and soft attractive potentials, and developed an erf 
%model for spheres with steep, short-range soft repulsive 
%potentials~\cite{schmidt2000fluid}.
% like that exhibited by a WCA fluid. 
%
Results will be compared to those from Monte-Carlo simulations.

@@

Another fairly simple, but relatively accurate interatomic potential widely 
used to represent generic, neutral atoms is the Leonard-Jones potential. 
The Leonard-Jones potential describes both the long-range attraction, 
and the short-range replusion that occurs between two nuetral atoms. 
The attraction between two nuetral atoms is due to fleeting dipoles that 
form spontaneously when electrons are redistributed as atoms come into 
close proximty. These are often characterized as Van der Walls, or London 
dispersion forces. The attraction grows stronger as two atoms move closer 
together until their atomic orbitals begin to overlap. 
%At this point the electrons in each atom begin to repel the electrons 
%in the other atom and oppose the overlap of the two atoms.
At this point the electrons in each atom begin to oppose the overlap of 
the two atoms through their replusion. The repulsion increases rapidly and 
approaches infinity as the distance between the atoms conintues to decrease.

@@

Freezing is governed by entropy, which the universe, or any closed system, 
seeks to maximize according to the second law of Thermodynamics. When the 
entropy of a pure substance in its liquid state is lower than the entropy 
that it would have in its solid state at the same temperature (necessary 
for thermal equilibrium), the same pressure (necessary for mechanical 
equilibrium), and the same chemical potential (necessary for diffusive 
equilibrium), an abrupt phase transition occurs, that is, the liquid freezes 
forming a crystalline structure. The particular crystalline structure and 
density that result are affected by the interatomic potential. When repulsive, 
soft sphere fluids (that are not too soft) freeze, they form lattice 
structures composed of Face Centered Cubic (FCC) crystal cells~\cite{Hansen}, 
as do hard spheres. 
%This structure is the three-dimensional equivilant to the two-dimensional 
%structure that one would see by filling the bottom of a box with a single layer of marbles. 
This structure is what one would see by filling a cube with marbles. It 
results in the highest density achievable for hard, and nearly-hard, spheres 
that are purely repulsive.

For pure substances, a phase transition occurs along a line on a 
Pressure-Temperature (P-T) phase diagram, such as can be seen in the phase 
diagram shown in Figure \ref{fig:P-T_Diagram}. This phase diagram is 
representative of some arbitrary, unknown substance composed of atoms, 
or molecules which exibit both attraction and repulsion. However, since 
the atoms in a WCA fliud do not exhibit attraction there will not be a 
dividing line between the gas and liquid phases, and so the term ``liquid" 
or ``gas" is not used, but simply ``fluid". Points along the line that 
separates the solid and the fluid phases indicate a phase transition, 
and show that the solid and the fluid both have the same temperature and 
the same pressure at a phase transistion.

@@

%In a process called Maxwell's Double Tangent construction, two curves are 
%used to find the densities of the liquid and the solid at the point of transistion. 
%Figure \ref{fig:MaxwellDT} shows a curve for the liquid, and a curve for 
%the solid both at a fixed temperature. Since the pressure of the liquid 
%is the same as the pressure of the solid at the point of transistion, the 
%density of the liquid and the density of the solid at the point of freezing 
%can be found from identifying the inverse densties at which the tangents 
%to each curve have the same slope.%
Two such curves are shown in Figure \ref{fig:MaxwellDT}, one for the liquid 
and one for the solid at a fixed temperature. Since the pressure of the 
liquid is the same as the pressure of the solid at the point of transistion, 
the density of the liquid and the density of the solid at the point of 
freezing can be found from identifying the inverse densties at which the 
tangents to each curve have the same slope. This process is called 
Maxwell's Double Tangent Construction.

@@

The minimization can be illustrated by considering that according to the 
second law of thermodynamics, a system that is not in equilibrium will 
exprience spontanteous processes that increase the total entropy of a 
system and its surroundings until equilibrium is reached, 
$\Delta{S}_{System + Surroundings} \geq 0$. Here ``surroundings" means 
the entire universe minus the system. This, together with the conservation 
of energy (the first law of thermodynamics) which says that energy lost 
(or gained) by the system is gained (or lost) by the surroundings 
$\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that 
changes in the free energy will be negative for spontaneous processes 
when the natural variables of the free energy are held fixed~\cite{schroeder}. 
%And so the free energy tends toward a minimum. 
%How will a system at a fixed temperature, volume, and number of 
%particles respond to a small offset from equilibrium? 
%
%Save for me:
%Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
%\begin{equation}
  %df_{N-System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\left(\frac{S}{N}\right)dT_{System}
%\end{equation}
%and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that 
%$dT=0$ gives
%\begin{equation}
  %f_{N-System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}
%\end{equation}
%where conservation of energy was also applied. Fixing the volume and the 
%number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, 
%results in
%\begin{equation}
  %df_{N-System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings}
%\end{equation}
%which, for fixed N, is the same as
%\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} 
%Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number 
%of particles are fixed, the change in the free energy of the system will 
%always be negative, and thus the free energy tends toward a 
%minimum~\cite{schroeder}. This illustrates the tendancy toward minimization 
%for a system near equilibrium. 
%End save for me.
A general proof of minimization 
can be found in Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}.

@@

Now consider the excess free energy which is written most generally as
\begin{equation}\label{Fexcess-mostgeneral}{F_{excess}= -k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)-V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\right)}}\end{equation} 
This integral can be greatly simplified if it can be assumed that the 
potential energy $V(\vec{r_1},\vec{r_2},...\vec{r_N})$ is a sum of the 
potential energies between pairs of atoms, where each atom is paired in 
turn with each of the other atoms in the system. In this case $V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ 
becomes 
\begin{equation}{V(\vec{r_1},\vec{r_2},...\vec{r_N})=\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}\end{equation} 
%\begin{displaymath}{=V_{12}(\vec{r_1},\vec{r_2})+V_{13}(\vec{r_1},\vec{r_3}) + V_{14}(\vec{r_1},\vec{r_4}) + ...}\end{displaymath}
%\begin{displaymath}{+ V_{23}(\vec{r_2},\vec{r_3})+V_{24}(\vec{r_2},\vec{r_4})+V_{25}(\vec{r_2},\vec{r_5})+ ...}\end{displaymath} 
%\begin{displaymath}{+ V_{34}(\vec{r_3},\vec{r_4})+V_{35}(\vec{r_3},\vec{r_5})+V_{36}(\vec{r_3},\vec{r_6})+...}\end{displaymath}
Using this potential, and setting the external potential 
$V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ to zero, the exponent in 
Equation~\ref{Fexcess} becomes

@@

%\begin{displaymath}F_{homogeneous} = F_{ideal.homogeneous} + F_{excess.homogeneous}\end{displaymath}
%\begin{displaymath}F_{crystal} = F_{ideal.crystal} + F_{excess.crystal}\end{displaymath}
%\newline
%\newline\noindent Inputs: 
%t*, n*, range of gw values , range of fv values
%\newline Outputs: Homogeneous (liquid)  Helmholtz free energy (at the given T,V,N) 
%\newline\indent\indent\indent Crystal Helmholtz free energy (at the given T,V,N)

%PROGRAM FLOWCHART ---------------------------
%\begin{displaymath}\text{Input:~~~~~~~~~~~~~~~~~~~~~~~~~~~T,~~ n,~~Range of fv,~~Range of gw~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath} 
%\begin{displaymath}\text{Set:~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~N~=~1~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath} 
%\begin{displaymath}\text{Best}\left(\frac{F}{N}\right)_{diff}= 1e100\end{displaymath}
%Loop: 
%\begin{displaymath}\text{fv = one of the fv values from input}\end{displaymath}
%\begin{equation}N_{primitive.cell}=(1-\text{fv})\text{N}\end{equation}
%\begin{equation}V_{primitive.cell}=\frac{\text{1}}{\text{n}}(1-\text{fv})\text{N}\end{equation}
%\begin{displaymath}\sigma = \text{one of the gw values from input}\end{displaymath}
%\begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int\exp^{-\frac{|\vec{r}-\vec{R}|^2}{2\sigma^2}}d\vec{r}}\end{equation} 
%\begin{equation}\left(\frac{F}{N}\right)_{crystal}=\frac{cFideal_{primitive.cell} + cFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
%\begin{equation}\left(\frac{F}{N}\right)_{homogeneous}=\frac{hFideal_{primitive.cell} + hFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
%\begin{equation}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{crystal}-{~~}\left(\frac{F}{N}\right)_{homogeneous}\end{equation}
%\begin{displaymath}\text{If~~~~~~}\left(\frac{F}{N}\right)_{diff}<\text{~~~~Best}\left(\frac{F}{N}\right)_{diff}\end{displaymath}
%\begin{displaymath}\text{Then~~~~~~}\text{Best}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{diff}\end{displaymath}
%\begin{displaymath}\text{Output:~~~~~~~~~~~~ Best}\left(\frac{F}{N}\right)_{crystal}\text{~~and~~~~~~}\left(\frac{F}{N}\right)_{homogeneous}\text{~~~~for the given T, n ~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath}

\section{Computing the Liquid and Crystal Helmholtz free energies}
%-----------Comuting the Liquid and Crystal free energies ------

@@

\begin{equation}{n_i(\vec{r})= \int_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation} 
implemented computationally in the form:
\begin{equation}{n_i(\vec{r})= \sum_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)dV}}\end{equation} 
becomes:
\begin{equation}{n_i(\vec{r})= \sum_{NumPoints}(1-f_v)P_{\vec{dr}}w_i(|\vec{r}-\vec{r}'|)}\end{equation} 
where NumPoints is the number of vectors $\vec{r'}$ randomly generated, and $n_g(\vec{r'})$ 
(the value of the Gaussian at at radius of $|\vec{r'}|$) is equal to a probability density 
which when multiplied by dV gives the probability $P_{\vec{dr}}$ at which the value of 
$w_i(|\vec{r}-\vec{r}'|)$ will occur. 
Using the relation:
\begin{displaymath}{f_{average}=\sum_i^N{P_if_i}=\sum_i{\frac{f_i}{N}}}\end{displaymath} 
$n_i(\vec{r})$ can be expressed as the average value of $(1-f_v)w_i(|\vec{r}-\vec{r}'|)$:
\begin{equation}{n_i(\vec{r})=\sum_{NumPoints}\frac{(1-f_v)w_i(|\vec{r}-\vec{r}'|)}{NumPoints}}\end{equation}
The result can be made more accurate by increasing NumPoints which is the 
number of vectors randomly generated. 
%-------Antithetic-------------------
%In order to further reduce the error 
%in the average, antithetic variates are used. Rather than computing an average 
%of a set of random points, each random point in the set is paired with a 
%point along the same line as the first point, but opposite in direction, 
%as shown in Figure \ref{fig:AntitheticVariate}. This reduces first order 
%error in the average by filtering out an erroneous offset of values along 
%the line connecting the two points by averaging the two results, one of 
%which may be erroneously higher than it should be while the other is 
%erronesously lower than it should be.

 %\begin{figure}[h!]
    %\centering
    %\includegraphics[height=4cm]{AntitheticVariate.png}
    %\caption{Antithetic variates reduce first order error in an average. 
    %Each point in the average is paired with another point from along 
    %the same line as the first, and at an equal, but opposite distance.}
    %\label{fig:AntitheticVariate}
  %\end{figure} 
%Incorporating antithetic variates $-\vec{dr}$ for every $\vec{dr}$ equation ?? becomes:
%\begin{displaymath}{n_i(\vec{r}_k, \vec{R}_j)=\sum_{NumPoints}\frac{1}{2}\frac{(1-f_v)\left[w_i(|\vec{r}-\vec{R_j}-\vec{dr}|) +w_i(|\vec{r}-\vec{R_j}+\vec{dr}|)\right]}{NumPoints}}\end{displaymath}   The process is repeated for all the other $\vec{R}_j$'s to get the total weighted densities at vector $\vec{r}_k$ due to all the Gaussians. 
%\begin{displaymath}{n_i(\vec{r}_k)=\sum_j{n_i(\vec{r}_k, \vec{R}_j)}}\end{displaymath}
%Then the entire process repeated for each vector $\vec{r}_k$ to get the 
%total weighted densities $n_i(\vec{r})$ as a function of the position $\vec{r}$.

%The standard error of the mean (SEM) is given by:
%\begin{equation}{SD_{mean}=\frac{SD}{\sqrt{N}}}\end{equation} which gives 
%a measure of the error in $n_3=\overline{w}_3$:
%\begin{equation}{Error=\sqrt{\frac{\overline{w_3^2}-(\overline{w}_3)^2}{NumPoints}}}\end{equation} 
%NumPoints, the number of vectors randomly generated, is increased until:
%\begin{equation}{Error<\frac{|1-n_3|}{4}}\end{equation} where the maximum 
%value that $n_3$ can be without error is 1.

%Figure \ref{fig:GaussandW2_actual} ......

@@

 \begin{figure}[h!]
    \centering
    \includegraphics[height=6cm]{GaussandW2_thinkas.png}
    \caption{An alternative way of viewing Figure \ref{fig:GaussandW2_actual} where 
    the weight function, shown by the purple dashed lines, is centered at $\vec{r}$  
    and fixed in place as $\vec{dr}$ varies randomly.} 
  \label{fig:GaussandW2_thinkas}
  \end{figure}

@@

Phase diagrams for the WCA fluid are shown in Figures~\ref{fig:Phase_Diagram_of_T_vs_n} 
and \ref{fig:Phase_Diagram_P_vs_T}. In all cases $\sigma=1$ and $\epsilon=1$. 
In Figure~\ref{fig:Phase_Diagram_of_T_vs_n} the reduced temperature $T^*=\frac{kT}{\epsilon}$ 
is plotted against the reduced number density $n^*=n\sigma^3$ for temperatures of kT = 0.5 to 3.0. 
The red area is fluid, and the blue area is crystal. The gray area that separates the fluid, 
and the crystal regions is where some fluid and crystal exist together at varying porportions. 
The region of coexistence slants slightly to the right toward higher density as the temperature 
increases. This is expected since at higher temperatures, the fluid would have to be under 
greater compression before forming a solid. The white lines show results from Monte-Carlo 
simulations of the WCA fluid. Overall, the general shape, and position of the lines correspond
fairly closely to the edges of the region of coexistance predicted from the DFT method, 
but the width of the coexistance region is about one-third larger for the Monte-Carlo results 
for most temperatures. The white line on the right, representing the density of the solid at 
each temperature, is in fairly good aggreement with our DFT results for all temperatures. 
However, the white line on the left, representing the density of the liquid at each temperature, 
is consistenty farther left than the liquid transition densities our method predicted.

@@

  %\cite{TschoppBHextended}

%\cite{rowlinson1964statistical}
%\cite{gil-villegas-1997-SAFT-VR, clark2006developing,
%  lafitte2013accurate}
%\cite{carnahan1969equation}
%\cite{rosenfeld1989, rosenfeld1997,
%  roth2002whitebear}

	%\cite{cuesta1997dimensional, hansen2009fundamental, marechal2013density, hughes2013classical, krebs2014improved}
%\cite{cuesta1997dimensional, hansen2009fundamental, marechal2013density}
%\cite{schmidt2000density, groh2001density, kim2001adsorption,
%  sweatman2002fundamental}

%\cite{weeks1971}
	%\cite{rosenfeld1989}
	%%\cite{tarazonaFIXME,santos2012phi3}
	%\cite{schmidt1999density} 
	%\cite{schmidt2000fluid}

%\cite{barker1967perturbation}
%\cite{roth2002whitebear} 
	%\cite{roth2002whitebear}

%\cite{mikolaj2004structure, eggert2002quantitative,
%yarnell1973structure}
	%\cite{verlet1967computer}

>>>>

The goal of this project is to generate computationally temperature-density 
and pressure-temperature phase diagrams for a Weeks-Chandler-Anderson (WCA) 
model fluid using classical Density Functional Theory (cDFT). 
To do this we need to find the Helmholtz free energies that the fluid
would have in a liquid state, and in a crystalline solid state 
at a given temperature and density. 
In general, the Helmholtz free energy, which consists of an ideal part 
and an excess part, can be written as a functional of the number density 
profile $n(\vec{r})$ which describes the way the average number 
density varies spatially. 
\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}

Classical Density Functional Theory is used to find the equilibrium 
Helmholtz free energy, and Soft Fundamental Measure Theory (SFMT) will 
be used to form the functional for the excess Helmholtz free energy. 
According to classical Density Functional Theory the equilbrium 
Helmholtz free energy of the system can be found by varying the number 
density profile $n(\vec{r})$ until the free energy is minimized. 
SFMT makes the task of forming an expression
for the excess Helmholtz free energy of a particular model fluid 
much easier since the interaction that each sphere, or atom, 
has with every other sphere does not have to be taken explicitly into account.
Results will be compared to those from Monte-Carlo simulations.

@@

Another fairly simple, but relatively accurate interatomic potential widely 
used to represent generic, neutral atoms is the Leonard-Jones potential. 
The Leonard-Jones potential describes both the long-range attraction, 
and the short-range replusion that occurs between two nuetral atoms. 
The attraction between two nuetral atoms is due to fleeting dipoles that 
form spontaneously when electrons are redistributed as atoms come into 
close proximty. These are often characterized as Van der Walls, or London 
dispersion forces. The attraction grows stronger as two atoms move closer 
together until their atomic orbitals begin to overlap. 
At this point the electrons in each atom begin to oppose the overlap of 
the two atoms through their replusion. The repulsion increases rapidly and 
approaches infinity as the distance between the atoms conintues to decrease.

@@

Freezing is governed by entropy, which the universe, or any closed system, 
seeks to maximize according to the second law of Thermodynamics. When the 
entropy of a pure substance in its liquid state is lower than the entropy 
that it would have in its solid state at the same temperature (necessary 
for thermal equilibrium), the same pressure (necessary for mechanical 
equilibrium), and the same chemical potential (necessary for diffusive 
equilibrium), an abrupt phase transition occurs, that is, the liquid freezes 
forming a crystalline structure. The particular crystalline structure and 
density that result are affected by the interatomic potential. When repulsive, 
soft sphere fluids (that are not too soft) freeze, they form lattice 
structures composed of Face Centered Cubic (FCC) crystal cells~\cite{Hansen}, 
as do hard spheres. 
This structure is what one would see by filling a cube with marbles. It 
results in the highest density achievable for hard, and nearly-hard, spheres 
that are purely repulsive.

For pure substances, a phase transition occurs along a line on a 
Pressure-Temperature (P-T) phase diagram, such as can be seen in the phase 
diagram shown in Figure~\ref{fig:P-T_Diagram}. This phase diagram is 
representative of some arbitrary, unknown substance composed of atoms, 
or molecules which exibit both attraction and repulsion. However, since 
the atoms in a WCA fliud do not exhibit attraction there will not be a 
dividing line between the gas and liquid phases, and so the term "liquid" 
or "gas" is not used, but simply "fluid". Points along the line that 
separates the solid and the fluid phases indicate a phase transition, 
and show that the solid and the fluid both have the same temperature and 
the same pressure at a phase transistion.

@@

%In a process called Maxwell's Double Tangent construction, two curves are 
%used to find the densities of the liquid and the solid at the point of transistion. 
%Figure \ref{fig:MaxwellDT} shows a curve for the liquid, and a curve for 
%the solid both at a fixed temperature. Since the pressure of the liquid 
%is the same as the pressure of the solid at the point of transistion, the 
%density of the liquid and the density of the solid at the point of freezing 
%can be found from identifying the inverse densties at which the tangents 
%to each curve have the same slope.%
Two such curves are shown in Figure~\ref{fig:MaxwellDT}, one for the liquid 
and one for the solid at a fixed temperature. Since the pressure of the 
liquid is the same as the pressure of the solid at the point of transistion, 
the density of the liquid and the density of the solid at the point of 
freezing can be found from identifying the inverse densties at which the 
tangents to each curve have the same slope. This process is called 
Maxwell's Double Tangent Construction.

@@

The minimization can be illustrated by considering that according to the 
second law of thermodynamics, a system that is not in equilibrium will 
exprience spontanteous processes that increase the total entropy of a 
system and its surroundings until equilibrium is reached, 
$\Delta{S}_{System + Surroundings} \geq 0$. Here ``surroundings" means 
the entire universe minus the system. This, together with the conservation 
of energy (the first law of thermodynamics) which says that energy lost 
(or gained) by the system is gained (or lost) by the surroundings 
$\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that 
changes in the free energy will be negative for spontaneous processes 
when the natural variables of the free energy are held fixed~\cite{schroeder}. 
A general proof of minimization 
can be found in Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}.

@@

Now consider the excess free energy which is written most generally as
\begin{equation}\label{Fexcess-mostgeneral}{F_{excess}= -k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)-V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\right)}}\end{equation} 
This integral can be greatly simplified if it can be assumed that the 
potential energy $V(\vec{r_1},\vec{r_2},...\vec{r_N})$ is a sum of the 
potential energies between pairs of atoms, where each atom is paired in 
turn with each of the other atoms in the system. In this case 
$V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ becomes 
\begin{equation}{V(\vec{r_1},\vec{r_2},...\vec{r_N})=\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}\end{equation} 
%\begin{displaymath}{=V_{12}(\vec{r_1},\vec{r_2})+V_{13}(\vec{r_1},\vec{r_3}) + V_{14}(\vec{r_1},\vec{r_4}) + ...}\end{displaymath}
%\begin{displaymath}{+ V_{23}(\vec{r_2},\vec{r_3})+V_{24}(\vec{r_2},\vec{r_4})+V_{25}(\vec{r_2},\vec{r_5})+ ...}\end{displaymath} 
%\begin{displaymath}{+ V_{34}(\vec{r_3},\vec{r_4})+V_{35}(\vec{r_3},\vec{r_5})+V_{36}(\vec{r_3},\vec{r_6})+...}\end{displaymath}
Using this potential, and setting the external potential 
$V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ to zero, the exponent in 
Equation~\ref{Fexcess-mostgeneral} becomes

@@

\section{Computing the Liquid and Crystal Helmholtz free energies}

@@

\begin{equation}{n_i(\vec{r})= \int_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation} 
implemented computationally in the form:
\begin{equation}{n_i(\vec{r})= \sum_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)dV}}\end{equation} 
becomes:
\begin{equation}{n_i(\vec{r})= \sum_{NumPoints}(1-f_v)P_{\vec{dr}}w_i(|\vec{r}-\vec{r}'|)}\end{equation} 
where NumPoints is the number of vectors $\vec{r'}$ randomly generated, and $n_g(\vec{r'})$ 
(the value of the Gaussian at at radius of $|\vec{r'}|$) is equal to a probability density 
which when multiplied by dV gives the probability $P_{\vec{dr}}$ at which the value of 
$w_i(|\vec{r}-\vec{r}'|)$ will occur. 
Using the relation:
\begin{displaymath}{f_{average}=\sum_i^N{P_if_i}=\sum_i{\frac{f_i}{N}}}\end{displaymath} 
$n_i(\vec{r})$ can be expressed as the average value of $(1-f_v)w_i(|\vec{r}-\vec{r}'|)$:
\begin{equation}{n_i(\vec{r})=\sum_{NumPoints}\frac{(1-f_v)w_i(|\vec{r}-\vec{r}'|)}{NumPoints}}\end{equation}
The result can be made more accurate by increasing NumPoints which is the 
number of vectors randomly generated.

@@

 \begin{figure}[h!]
    \centering
    \includegraphics[height=6cm]{GaussandW2_thinkas.png}
    \caption{An alternative way of viewing Figure~\ref{fig:GaussandW2_actual} where 
    the weight function, shown by the purple dashed lines, is centered at $\vec{r}$  
    and fixed in place as $\vec{dr}$ varies randomly.} 
  \label{fig:GaussandW2_thinkas}
  \end{figure}

@@

Phase diagrams for the WCA fluid are shown in Figures~\ref{fig:Phase_Diagram_of_T_vs_n} 
and \ref{fig:Phase_Diagram_P_vs_T}. In all cases $\sigma=1$ and $\epsilon=1$. 
In Figure~\ref{fig:Phase_Diagram_of_T_vs_n} the reduced temperature $T^*=\frac{kT}{\epsilon}$ 
is plotted against the reduced number density $n^*=n\sigma^3$ for temperatures of kT = 0.5 to 3.0. 
The red area is fluid, and the blue area is crystal. The gray area that separates the fluid, 
and the crystal regions is where some fluid and crystal exist together at varying porportions. 
The region of coexistence slants slightly to the right toward higher density as the temperature 
increases. This is expected since at higher temperatures, the fluid would have to be under 
greater compression before forming a solid. The white lines show results from Monte-Carlo 
simulations of the WCA fluid~\cite{May}. Overall, the general shape, and position of the lines correspond
fairly closely to the edges of the region of coexistance predicted from the DFT method, 
but the width of the coexistance region is about one-third larger for the Monte-Carlo results 
for most temperatures. The white line on the right, representing the density of the solid at 
each temperature, is in fairly good aggreement with our DFT results for all temperatures. 
However, the white line on the left, representing the density of the liquid at each temperature, 
is consistenty farther left than the liquid transition densities our method predicted.

@@
